THESIS TITLE

### 2023-07-11 msthesis.tex
simulation chapter half way done

\documentclass[a4paper,oneside,12pt]{report}
\usepackage{styles/fbe_tez}
\usepackage[utf8x]{inputenc} % To use Unicode (e.g. Turkish) characters
\usepackage{amsmath, amsthm} % Some extra symbols
\usepackage[bottom]{footmisc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{amssymb,bm,cite,graphicx, fixmath, texdraw}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{notoccite}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{minted}

@@

\subsection{Rhodotron Accelerator}

@@

\chapter{SIMULATION}
\newpage

@@

@@

@@

>>>>

\documentclass[a4paper,oneside,12pt]{report}
\usepackage{styles/fbe_tez}
\usepackage[utf8x]{inputenc} % To use Unicode (e.g. Turkish) characters
\usepackage{amsmath, amsthm} % Some extra symbols
\usepackage[bottom]{footmisc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{amssymb,bm,cite,graphicx, fixmath, texdraw}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{notoccite}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{minted}
\usepackage{xcolor}

@@

\subsection{Rhodotron Accelerator} \label{sec:theory_rhodo}

@@

\chapter{SIMULATION}
% Start of poc.tex
\section{Proof of Concept}
Once the need for a custom built tool for designing and simulating a Rhodotron type accelerator was apparent, a simple 1D proof of concept was implemented.

This proof of concept worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}

The implemenation of this logic can be observed in the following code block.
\begin{figure}[H]
    \begin{minted}[linenos=true]{c++}
        for (double t=0; t<SimuTime; t+=dT){
            double RelBeta  = v/c;
            double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

            double ef=Eradial(r_pos*1000,t,0); // convert position to mm
            double acc=ef*1E6*eQMratio/(RelGamma*RelGamma*RelGamma);

            r_pos = r_pos + v * dT*ns + 1/2*acc*(dT*ns)*(dT*ns);
            v = v + acc*dT*ns;

            RelBeta  = v/c;
            RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);
            Et=RelGamma*E0; 
        }
    \end{minted}
    \caption{Core logic loop of the POC}
    \label{fig:POC_core_logic}
\end{figure}

After simulating the \textit{synchronous electron} mentioned in \fromsec{cavity_of_a_rhodotron} with $\phi_{lag}=15^\circ$ for one pass in THE CAVITY, the results from \textit{POC}, \textit{CST} and \fromeq{W_gain_each_pass_pottier} are compared.
% 40 kW total power
% Emax = 0.96 in old, 2.08 in new version
% R1 = 0.1882
% R2 = 0.753
% f = 107.5
\begin{eqnarray}
    E_{Pottier} &=& 0.565 MeV \nonumber\\
    E_{CST} &=& 0.872 MeV \label{eq:poc_E_results}\\
    E_{POC} &=& 0.873 MeV \nonumber
\end{eqnarray}

As can be observed in \fromeq{poc_E_results}, \textit{POC software} and \textit{CST} produced very similar results. However, we cannot make the same claim regarding \fromeq{W_gain_each_pass_pottier}.
This inconsistency between $E_{Pottier}$ and $E_{CST}$ was noticed in the earlier simulations as well, which supported the idea of using another simulation software to reduce reliance on $CST$.
Using a single software would have made the process error prone. Their results, at the end of the day, depend heavily on the physical and mathematical models they use.

After the promising results obtained from \textit{POC}, the decision was made to develop a more robust simulation software. This software was named \textit{Rhodotron Simulation} for the current stage and started development in October, 2021.
The development and improvement efforts are still ongoing.

During the development of \textit{Rhodotron Simulation} software, several intermediate builds have been implemented and tested. The following sections in this chapter investigates their implementations further.

@@

% Start of intermediate.tex
\section{Intermediate Versions}

\subsection{$L_{out}$ Optimization For Single $e^-$}
Initial step of improving $POC$ towards $Rhodotron Simulation$ was to implement \textit{$L_{out}$ optimizations} to help optimazing magnet designs, as discussed in \fromsec{parameter_sweep}.
First approach was to hang the $e^-$ outside of the cavity for $t_{out} = L_{out}/v$, then inject it back to the cavity with reversed $\vecthreeBF{v}$. Then sweep the $t_{out}$ parameter to find the optimal value.
This simple implementation can be found in \fromfig{lout_opt_single_e} of \fromapp{intermediate_codes}.

Although the results from this optimization sweep were promising after they were simulated with $CST$, simulating one particle would not be sufficiently useful for designing a magnet.

\subsection{$\phi_{lag}$ Optimization For Bunches}
After successfully accelerating single $e^-$, particle bunches were implemented to approximate a real $e^-$ gun. 
They were modeled as $N$ electrons fired from an $e^-$ gun at even time intervals. This approach was taken because the amount of time gun fires, defined as \textit{Gun Active Time, $t_g$}, is a crucial part of pulsing $e^-$ gun design.

Addition of bunches would immediately proven useful when finding optimal gun phase lag.
$\phi_{lag}$ for a bunch was defined as the RF phase when the first $e^-$ of the bunch entered the cavity, it defines the starting time of the current pass.
To use the \textit{parameter sweep} method, as used in \textit{$L_{out}$ optimization}, relevant bunch characteristics are defined as follows:

\begin{itemize}
    \item $\mu E$: Average energy
    \item $E_{RMS}$: Root mean square of energy
    \item $r_{RMS}$: Root mean square of $e^-$ positions
\end{itemize}

Optimal $\phi_{lag}$ would produce maximum $\mu E$ while minimizing $E_{RMS}$ \& $r_{RMS}$. For the first pass, 
$E_{RMS}$ and $r_{RMS}$ would be vaguely dependent of each other; therefore, early implementation of $\phi_{lag}$ sweep was based on minimizing $E_{RMS}$ during simulation (\textit{see \fromfig{phlag_opt}}). 
For $\mu E$ considerations, data from the software would be analyzed either manually or by using external tools such as $ROOT$.

\begin{figure}[H]
    \centering
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\small]{c++}
        int phase_opt(int phase_sweep_range){
            double minrms = 1;
            int opt_phase;
            for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
              Bunch bunch1(RFphase);
              double t1 = 0;
              bunch1.bunch_gecis_t(t1);
              bunch1.reset_pos();

              if( bunch1.E_rms() < minrms ){
                minrms = bunch1.E_rms();
                opt_phase = RFphase;
              }
            }
            return opt_phase;
        }
    \end{minted}
    \caption{$\phi_{lag}$ Optimization For Initial Bunch Design}
    \label{fig:phlag_opt}
\end{figure}

Since $\phi_{lag}$ is relatively easy to change after production, another version of \fromfig{phlag_opt} that was modified for given magnet design parameters was also implemented (\textit{see \fromfig{phlag_opt_n_pass} in \fromapp{intermediate_codes}}). 
This version can be useful for optimizing $\phi_{lag}$ in case of production issues in magnets.

\subsection{$L_{out}$ Optimization For Bunches}

% START OF APPENDIX

@@

\begin{figure}[H]
        \centering
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm]{c++}
            int phase_opt(const vector<double>& Louts, int phase_sweep_range){
                double minrms = 1;
                int opt_phase;
                for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
                    Bunch bunch1(RFphase);
                    double t1 = 0;
                    bunch1.bunch_gecis_t(t1);
                    bunch1.reset_pos();

                    for(int i = 0; i < Louts.size(); i++){
                        bunch1.bunch_gecis_d(Louts[i]);
                        bunch1.reset_pos();
                    }

                    if( bunch1.E_rms() < minrms ){
                        minrms = bunch1.E_rms();
                        opt_phase = RFphase;
                    }
                }
                return opt_phase;
            }
        \end{minted}
    \caption{$\phi_{lag}$ Optimization For Tnitial Bunch Design}
    \label{fig:phlag_opt_n_pass}
\end{figure}

% END OF APPENDIX

% START OF BIBLIOGRAPHY

@@

% END OF BIBLIOGRAPHY

### 2023-07-11 subsections/simulation/intermediate_versions/intermediate.tex
simulation chapter half way done

\subsection{From Single $e^-$ to Bunches}
After successfully accelerating single $e^-$, next step was to implement bunches to approximate a real $e^-$ gun.

@@

>>>>

\subsection{$\phi_{lag}$ Optimization For Bunches}
After successfully accelerating single $e^-$, particle bunches were implemented to approximate a real $e^-$ gun. 
They were modeled as $N$ electrons fired from an $e^-$ gun at even time intervals. This approach was taken because the amount of time gun fires, defined as \textit{Gun Active Time, $t_g$}, is a crucial part of pulsing $e^-$ gun design.

Addition of bunches would immediately proven useful when finding optimal gun phase lag.
$\phi_{lag}$ for a bunch was defined as the RF phase when the first $e^-$ of the bunch entered the cavity, it defines the starting time of the current pass.
To use the \textit{parameter sweep} method, as used in \textit{$L_{out}$ optimization}, relevant bunch characteristics are defined as follows:

\begin{itemize}
    \item $\mu E$: Average energy
    \item $E_{RMS}$: Root mean square of energy
    \item $r_{RMS}$: Root mean square of $e^-$ positions
\end{itemize}

Optimal $\phi_{lag}$ would produce maximum $\mu E$ while minimizing $E_{RMS}$ \& $r_{RMS}$. For the first pass, 
$E_{RMS}$ and $r_{RMS}$ would be vaguely dependent of each other; therefore, early implementation of $\phi_{lag}$ sweep was based on minimizing $E_{RMS}$ during simulation (\textit{see \fromfig{phlag_opt}}). 
For $\mu E$ considerations, data from the software would be analyzed either manually or by using external tools such as $ROOT$.

\begin{figure}[H]
    \centering
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\small]{c++}
        int phase_opt(int phase_sweep_range){
            double minrms = 1;
            int opt_phase;
            for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
              Bunch bunch1(RFphase);
              double t1 = 0;
              bunch1.bunch_gecis_t(t1);
              bunch1.reset_pos();

              if( bunch1.E_rms() < minrms ){
                minrms = bunch1.E_rms();
                opt_phase = RFphase;
              }
            }
            return opt_phase;
        }
    \end{minted}
    \caption{$\phi_{lag}$ Optimization For Initial Bunch Design}
    \label{fig:phlag_opt}
\end{figure}

Since $\phi_{lag}$ is relatively easy to change after production, another version of \fromfig{phlag_opt} that was modified for given magnet design parameters was also implemented (\textit{see \fromfig{phlag_opt_n_pass} in \fromapp{intermediate_codes}}). 
This version can be useful for optimizing $\phi_{lag}$ in case of production issues in magnets.

\subsection{$L_{out}$ Optimization For Bunches}

@@

\begin{figure}[H]
        \centering
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\small]{c++}
            int phase_opt(const vector<double>& Louts, int phase_sweep_range){
                double minrms = 1;
                int opt_phase;
                for(int RFphase = -phase_sweep_range; RFphase <= phase_sweep_range; RFphase++){
                    Bunch bunch1(RFphase);
                    double t1 = 0;
                    bunch1.bunch_gecis_t(t1);
                    bunch1.reset_pos();

                    for(int i = 0; i < Louts.size(); i++){
                        bunch1.bunch_gecis_d(Louts[i]);
                        bunch1.reset_pos();
                    }

                    if( bunch1.E_rms() < minrms ){
                        minrms = bunch1.E_rms();
                        opt_phase = RFphase;
                    }
                }
                return opt_phase;
            }
        \end{minted}
    \caption{$\phi_{lag}$ Optimization For Tnitial Bunch Design}
    \label{fig:phlag_opt_n_pass}
\end{figure}

### 2023-07-11 subsections/simulation/poc/poc.tex
simulation chapter half way done

This proof of concept worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_y}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}

@@

As can be observed in \fromfig{poc_E_results}, \textit{POC software} and \textit{CST} produced very similar results.However, we cannot make the same claim regarding \fromeq{W_gain_each_pass_pottier}.
This inconsistency between $E_{Pottier}$ and $E_{CST}$ was noticed in the earlier simulations as well, which supported the idea of using another simulation software to reduce reliance on $CST$.
Using a single software would make the process error prone. Their results, at the end of the day, depend on the physical and mathematical models they use.

>>>>

This proof of concept worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}

@@

As can be observed in \fromeq{poc_E_results}, \textit{POC software} and \textit{CST} produced very similar results. However, we cannot make the same claim regarding \fromeq{W_gain_each_pass_pottier}.
This inconsistency between $E_{Pottier}$ and $E_{CST}$ was noticed in the earlier simulations as well, which supported the idea of using another simulation software to reduce reliance on $CST$.
Using a single software would have made the process error prone. Their results, at the end of the day, depend heavily on the physical and mathematical models they use.
